---
title: "Lab2_Visualization"
author: "Anubhav Dikshit and Yusur Almutair"
date: "13 September 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Lab 2 loading library

```{r}
library(data.table)
library(ggplot2)
library(dplyr)
library(gridExtra)
library(plotly)
```

### Reading input 
```{r}
setwd("C:/Users/anubh/OneDrive/Documents/GitHub/LiU/LiU-2018/Visualization/Assignment_2/")

olive_data <- read.csv("Input/olive.csv")
olive_data <- olive_data[,-1]
```

### 1. Scatter plot dependence of Palmitic on Oleic, linoleic not classed and classfed
```{r}
ggplot(data = olive_data, aes(x = oleic, y = palmitic), colour = linoleic) + geom_point(aes(colour = linoleic))
ggplot(data = olive_data, aes(x = oleic, y = palmitic)) + geom_point(aes(colour = cut_interval(linoleic, n = 4)))
```
Analysis: The second graph is easier to analysis than compared to first one, the perception
problem highlighted here is the difference in the channel capacity of human preception. Where capacity of distingushing hue > intensity 

### 2. Scatter plot dependence of Palmitic on Oleic, linoleic split by different attributes
```{r}
ggplot(data = olive_data, aes(x = oleic, y = palmitic)) + geom_point(aes(colour = cut_interval(linoleic, n = 4)))
ggplot(data = olive_data, aes(x = oleic, y = palmitic), colour = linoleic) + geom_point(aes(size = cut_interval(linoleic, n = 4)))
ggplot(data = olive_data, aes(x = oleic, y = palmitic), colour = linoleic) + geom_point() +  geom_spoke(aes(angle = as.numeric(cut_interval(linoleic, n = 4))), radius = 1.0)
```
