---
title: "Advanced Data Mining Association Analysis-1"
author: "Nahid Farazmand (nahfa911) and Anubhav Dikshit(anudi287)"
date: "February 27, 2019"
output: pdf_document
---
In this exercise we want to cluster a given dataset and use association analysis to describe the clusters obtained. for this purpose, we will work with one of the most well-known datasets in the data mining literature, namely the Iris dataset. The dataset consists of 50 samples from each of three species of Iris flowers (Iris setosa, Iris virginica and Iris versicolor). Four features were measured from each sample, they are the length and the width of sepal and petal.

## First step: Clustering (by usin simple K-mean algorithm)
Here we know there are 3 types of Iris flowers; so K is 3 and we have chosen the seed = 10 (from the previous exercise we know that changing the rate of seed will change the initial starting points of the K-mean process). By using K-mean the result will be as follow:

```{r echo=FALSE, message=FALSE, warning=FALSE}
library(foreign)
data <- read.arff('iris2.arff')
data[order(data$cluster),c(5,6)]
``` 

\includegraphics[]{1-1.Kmeans.PNG}

\includegraphics[]{1-3.Kmeans-clustersvsclass.PNG}

As we can see, cluster 1 is exactly same as class 1, cluster 1 contains two observations from class 3 and cluster 2 contains 7 observations from class 2.

## Second step: Association analysis (Apriori algorithm)
Here, at first we create a new attribute that represents the cluster label assigned to each instance.

After that we will implement Apriori algorithm and then we try to Find rules that are accurate and such that the antecedent does not contain the class attribute and the consequent only contains the cluster attribute. Such rules for the 3 clusters should help us to describe the instances grouped in each cluster.

### 1. Number of clusters = 3 (in clustering),NUmber of bins = 3 (In discretizing), Minimum support: 0.3 (45 instances), Minimum metric (confidence): 0.8
Founded rules:

1. sepallength='(-inf-5.5]' petallength='(-inf-2.966667]' 47 ==> cluster=cluster3 47  

2. sepallength='(-inf-5.5]' petalwidth='(-inf-0.9]' 47 ==> cluster=cluster3 47 

3. sepallength='(-inf-5.5]' petallength='(-inf-2.966667]' petalwidth='(-inf-0.9]' 47 ==> cluster=cluster3 47 

### 1. Number of clusters = 3 (in clustering),NUmber of bins = 2 (In discretizing), Minimum support: 0.4 (60 instances), Minimum metric (confidence): 0.8

```{r echo=FALSE, message=FALSE, warning=FALSE}
data2 <- read.arff('iris3.arff')
data2[order(data2$cluster),c(5,6)]
```

\includegraphics[]{2.1.Kmeans-clustersvsclass.PNG}

Founded rules:

1. sepallength='(-inf-6.1]' petalwidth='(-inf-1.3]' 74 ==> cluster=cluster1 74  

2. petallength='(3.95-inf)' petalwidth='(1.3-inf)' 71 ==> cluster=cluster3 71 

3. petallength='(-inf-3.95]' 61 ==> cluster=cluster1 61  

4. sepallength='(-inf-6.1]' petallength='(-inf-3.95]' 61 ==> cluster=cluster1 61  

5. sepalwidth='(-inf-3.2]' petallength='(3.95-inf)' petalwidth='(1.3-inf)' 61 ==> cluster=cluster3 61 

6. petallength='(-inf-3.95]' petalwidth='(-inf-1.3]' 60 ==> cluster=cluster1 60 

7. sepallength='(-inf-6.1]' petallength='(-inf-3.95]' petalwidth='(-inf-1.3]' 60 ==> cluster=cluster1 60  

8. petalwidth='(1.3-inf)' 72 ==> cluster=cluster3 71  

9. sepalwidth='(-inf-3.2]' petalwidth='(1.3-inf)' 62 ==> cluster=cluster3 61

10. petallength='(-inf-3.95]' 61 ==> petalwidth='(-inf-1.3]' cluster=cluster1 60   

11. petalwidth='(-inf-1.3]' 78 ==> cluster=cluster1 74  

### 1. Number of clusters = 4 (in clustering),NUmber of bins = 2 (In discretizing), 0.3 (45 instances), Minimum metric (confidence): 0.8
```{r echo=FALSE, message=FALSE, warning=FALSE}
data3 <- read.arff('iris4.arff')
data3[order(data3$cluster),c(5,6)]
```

\includegraphics[]{3.1.Kmeans-clustersvsclass.PNG}

Founded rules:

1. petallength='(-inf-3.95]' petalwidth='(-inf-1.3]' 60 ==> cluster=cluster4 60  

2. sepallength='(-inf-6.1]' petallength='(-inf-3.95]' petalwidth='(-inf-1.3]' 60 ==> cluster=cluster4 60  

3. sepallength='(6.1-inf)' petalwidth='(1.3-inf)' 51 ==> cluster=cluster3 51  

4. sepallength='(6.1-inf)' petallength='(3.95-inf)' petalwidth='(1.3-inf)' 51 ==> cluster=cluster3 51  

5. sepallength='(6.1-inf)' petalwidth='(1.3-inf)' 51 ==> petallength='(3.95-inf)' cluster=cluster3 51  


