---
title: "cstats_lab05"
author: "Thijs Quast (thiqu264), Anubhav Dikshit (Anudi287)"
date: "19-2-2019"
output: pdf_document
---
# Question 1 - Hypothesis testing
```{r}
library(readxl)
lottery <- read_xls("lottery.xls")
Y <- lottery$Draft_No
X <- lottery$Day_of_year
```

## 1.1
```{r}
library(ggplot2)
plot <- ggplot(lottery, aes(x = X, y = Y)) + geom_point()
plot <- plot + xlab("Day of the Year") + ylab("Draft number") + ggtitle("Draft number vs Day of the Year")
plot
```
The lottery looks random, no pattern between Draft number and Day of teh Year seems to be present.

## 1.2
```{r}
loess_model <- loess(Y ~ X, data = lottery)
Yfitted <- loess_model$fitted
plot <- plot + geom_line(aes(x = X, y =Yfitted, col = "Yfitted"))
plot
```
Still, largely the lottery looks random. However, if it would be perfectly random there would be an exact horizontal line, now there seems to be slight downward trend in the data. This implies that actually people who were born at the end of the year have a higher chance of getting a lower draft number and therefore being drafted earlier.

## 1.3
```{r}
library(boot)

stat1<-function(data,vn){
    # Data frame for the bootstrap
    data<-as.data.frame(data[vn,])
    # Model and prediction
    loess_model2 <- loess(Draft_No ~ Day_of_year, data)
    prediction <- predict(loess_model2, data)
    
    # Statistics for T
    X_a <- data$Day_of_year[which.max(data$Draft_No)]
    X_b <- data$Day_of_year[which.min(data$Draft_No)]
    Y_a <- prediction[X_a] 
    Y_b <- prediction[X_b]
    
    # Output
    t <- ((Y_b - Y_a) / (X_b - X_a))
    return(t)
    
}

bootstrap <- boot(lottery,stat1,R=2000)
```

```{r}
print(boot.ci(bootstrap))
plot(bootstrap)
```
```{r}
t_final <- mean(bootstrap$t)
t_final
p_value <- pt(t_final, df = (nrow(lottery)-1))
p_value
```
We get a p-value of 0.40, meaning we cannot reject the null hypothesis. Meaning we cannot say the lottery is not random.

## 1.4
```{r}
permutation <- function(data, B){
  
}
```

## 1.5

# Assignment 2 - Bootstrap, jackknife and confidence intervals
## 2.1
```{r}
prices <- read_excel("prices1.xls")
```

```{r}
hist(prices$Price, breaks = 50)
```
I would say this data shows the shape of a beta distribution.
```{r}
mean(prices$Price)
```

## 2.2
```{r}
library(boot)

stat2 <- function(data, vn){
  data <- as.data.frame(data[vn,])
  output <- mean(data$Price)
}

res <- boot(prices, stat2, R = 2000)
res
plot(res)
```

```{r}
var(res$t)
```

```{r}
boot.ci(res, type = "perc")
```

```{r}
boot.ci(res, type = "bca")
```
```{r}
boot.ci(res, type = "norm")
```

## 2.3
```{r}
library(bootstrap)
res2 <- jackknife(prices$Price, mean)
summary(res2)
res2
```

```{r}
confidence_interval_jackknife <- c(mean(res2$jack.values - 1.96*res2$jack.se), 
                                   mean(res2$jack.values + 1.96*res2$jack.se))
```

