---
title: "Visualization Lab 1 - Shiny App"
author: "Hariprasath Govindarajan (hargo729), Lakshidaa Saigiridharan (laksa656)"
date: "13 September 2018"
output: html_document
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Task 2.8 - Shiny App

Shiny application that allows the user to interactively enable/disable density plots for variables in SENIC data along with option to select value for bandwidth parameter in density estimation.

```{r, echo=FALSE, message=FALSE, warning=FALSE}

#-----------------------------------------------------------------------

#Task 2.8

#Shiny Application

# Loading required R packages
library(ggplot2)
library(plotly)
library(shiny)
library(gridExtra)

senic_data = read.table("SENIC.txt", 
                  col.names = c("Identification_Number", "Length_of_Stay", "Age",
                                "Infection_Risk", "Routine_Culturing_Ratio", 
                                "Routine_Chest_X_ray_Ratio", "Number_of_Beds",
                                "Medical_School_Affiliation", "Region", 
                                "Average_Daily_Census", "Number_of_Nurses",
                                "Available_Facilities_and_Services"))

get_outliers <- function(x){
  quantile_values = quantile(x, probs = c(0.25, 0.75))
  q1 = quantile_values["25%"]
  q3 = quantile_values["75%"]
  
  return(c(which((x > (q3+1.5*(q3-q1)))), which(x < (q1-1.5*(q3-q1)))))
}

categorical_columns = c("Medical_School_Affiliation", "Region")
ID_columns = c("Identification_Number")
quantitative_columns = setdiff(colnames(senic_data), c(categorical_columns, ID_columns))

#UI component
ui <- fluidPage(
  sliderInput(inputId="bw_value", label="Choose bandwidth size", value=4.5, min=0.1, max=80),
  checkboxGroupInput("selected_variables", "Variables to show: ", quantitative_columns, inline=TRUE),
  plotOutput("densPlot", height = "650px")
)

plot_density_with_outliers_shiny <- function(df_data, col_name, bw){
  p <- NULL
  if(length(get_outliers(senic_data[[col_name]])) > 0){
    p <- ggplot(df_data) + 
      ggtitle(paste("Density plot of ", col_name)) + 
      theme(plot.title = element_text(hjust = 0.5)) + 
      geom_density(aes_string(x=col_name), fill = "lightblue", color = "darkblue", bw=bw) +
      geom_point(data=df_data[get_outliers(df_data[[col_name]]),],
                 aes_string(x=col_name, y=0), shape=23, size=2, colour="black", fill="red")
  }
  else{
    p <- ggplot(df_data) + 
      ggtitle(paste("Density plot of ", col_name)) + 
      theme(plot.title = element_text(hjust = 0.5)) + 
      geom_density(aes_string(x=col_name), fill = "lightblue", color = "darkblue", bw=bw)
  }
  
  return(p)
}

# Server component
server <- function(input, output) {
  
  output$densPlot <- renderPlot({
    
    selected_columns = input$selected_variables
    plot_list = vector("list", length(selected_columns))
    
    if(length(selected_columns) > 0){
      for(i in 1:length(selected_columns)){
        plot_list[[i]] = plot_density_with_outliers_shiny(senic_data, selected_columns[i], 
                                                    bw = input$bw_value)
      }
      plot_matrix <- arrangeGrob(grobs = plot_list, ncol = 2)
      grid.arrange(plot_matrix)
    }
    
  })
}

# Run the application 
shinyApp(ui = ui, server = server, options = list(width="800px", height="900px"))

```

We observe that increasing the bandwidth makes the plots smoother. We cannot state an optimal value of bandwidth that applies for all the 9 density plots. Number of beds, Number of nurses and avarage daily census have much higher data values (order of 0-800) and range compared to the other variables and require a much higher optimal bandwidth (bandwidth = 65). At such a bandwidth, the other plots are reduced to almost straight lines. On the other hand, the infection risk variable has very low data values (order of 8) and range and requires a much lower bandwidth for optimal viewing (bandwidth = 0.5). At such a bandwidth many other plots are still chunky and not smooth. For the other variables, an optimal bandwidth can be set at 4.5. 

# Appendix
```{r, ref.label=knitr::all_labels(),echo=TRUE,eval=FALSE}
```